Preliminary Draft

### 2021-07-21 article/draft.tex
rewrite and reduce a bit the Fb reduce section

%Another measure that mainstream social media platforms can apply is to prevent users from sharing specific types of content, in this example URLs coming from a specific domain name.

@@

Mainstream Social Media platforms can reduce the visibility of the content created or shared by specific users, whenever they violate the platforms' rules. The implementation of this policy varies across platforms and is not easy to verify ex-post. In what follows we provide means to verify this policy on Twitter and Facebook.

@@

%phrase pas accès au reach, mais seulement les engagement metrics
% Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist\footnote{See \href{https://meta.wikimedia.org/wiki/Spam\_blacklist}{meta.wikimedia.org/wiki/Spam\_blacklist}.},
%CrowdTangle and Buzzsumo provide access to the interaction metrics on content (engagement), but not to the number of views (or reach), therefore we can only verify the {\it reduce} measure indirectly through a decrease in engagement. If Facebook was reducing the reach of a content via a lower ranking, we would expect the content to be less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

@@

On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless, they are explicitly condemning the material.\footnote{See the article in Wired by Martineau (2020)~\cite{wiredalexjones} and see the section {\it So what happened with InfoWars? They were up on Friday and now they are down?} \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{about.fb.com/news/2018/08/enforcing-our-community-standards/}.} To verify the measure, we used the ``/posts/search'' endpoint\footnote{see the documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Search}{github.com/CrowdTangle/API/wiki/Search}.} of the CrowdTangle API, to collect $37$ $242$ Facebook public posts that had shared a URL link containing ``infowars.com'', published between January $1$, $2019$ and December $31$, $2020$.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the $27$ $721$ posts directly sharing an Infowars link.} The command used can be found in the following Github repository: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_infowars_data.sh}{link}. The number of public posts sharing an $Infowars$ link remained globally stable throughout $2019$ (see figure \ref{infowars1} top panel). Thus the measure announced by Facebook doesn't seem to have prevented users from sharing an $Infowars$ link. Nevertheless, a clear drop in engagement was observed on May 2, 2019 (see figure \ref{infowars1} bottom panel). The number of reactions, shares and comments per post have decreased respectively  by $-94\%$,  $-96\%$ and $-93\%$  when comparing the two months before and after May 2, 2019. This suggests the measure taken by Facebook in May $2019$, is not a ban per se, but rather a $reduce$ measure. This is because users could still post $Infowars$ links, but these posts generated less engagement. It should be noted that the engagement metrics increased again by the end of $2019$ / beginning of $2020$, suggesting that the $reduce$ measure may have been lifted a few months after its implementation.

	%\begin{multicols}{1}
		\includegraphics[scale=0.35]{./img/infowars/fb_infowars_3.png}
	%\end{multicols}

	\caption{Facebook engagement metrics (average number of reactions, shares and comments per article) for the Infowars articles published in $2019-2020$ and gathered from the Buzzsumo API. The red line marks the date of May $2$, $2019$, when Facebook has announced the ban regarding Infowars.}
	\label{infowars2}
\end{figure}

As CrowdTangle is tracking posts only from certain public groups and pages, we also used the ``/search/articles'' endpoint of the Buzzsumo API, to gather a richer Facebook dataset. We collected the engagement data for the $14$ $232$ articles crawled by Buzzsumo from the $Infowars$ website between January $1$, $2019$ and December $31$, $2020$.  %\footnote{The command can be found: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_infowars\_data.py}{here}.} 
We observe that the articles published after May $2$, $2019$ received less Facebook engagement than the ones published before (see figure \ref{infowars2}), with a percentage change of -$97\%$ for the reactions, -$59\%$ for the shares and -$97\%$ for the comments. An increase in engagement was also observed in $2020$. It reinforces the hypothesis that Facebook reduced the reach of posts sharing Infowars links only during a few months in $2019$.

We found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes around 800 articles per month, only 53 articles were collected between June 11 to July 11, 2020\footnote{In Figure \ref{infowars2} the engagement metrics were only shown for days in which at least 5 Infowars articles were collected from Buzzsumo, to ensure reliable results.}. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

\subsubsection{Facebook}

The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} Facebook has announced on December $20$, $2019$ that {\it The BL is now banned from Facebook} for coordinated inauthentic behavior (see Facebook Newsroom (2019)~\cite{newsroom1}), which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. Coordinated inauthentic behavior is a distinct phenomenon from disinformation according to Facebook, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false”.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.} Nevertheless, for the case of the Beauty of Life, both misinformation and coordinated inauthentic behavior were attested, according to the fact-checking organization Snopes which had reported about The BL’s activity to Facebook and other various public articles\footnote{\href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}}.

@@

To verify Facebook’s ban of The BL domain name, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from the Buzzsumo API\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.}. We used the ``/search/articles'' endpoint to collect the engagement metrics of the $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.%\footnote{The command can be found in the following Github repository: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}.}

The number of Facebook reactions, shares and comments dropped to zero for TheBL’s articles published after December $1$, $2019$ (see figure \ref{fb_bl} top panel), indicating the start of the ban. We can note that although the ban was communicated in an article~\cite{newsroom1} published on December $20$, $2019$, it seems to have actually started on December $1$, $2019$.

The communication around the ban appeared to have discouraged The Beauty of Life to proceed with their activity. Indeed the number of articles they published daily was around $50$ until December $20$, $2019$, when it decreased drastically to reach around $5$ to $10$ articles published per day (see figure \ref{fb_bl} bottom panel). Using Buzzsumo data, we ascertained that links from thebl.com were not shared on Facebook anymore. The ban started on December $1$, $2019$, and appeared to be still enforced in June 2021.

>>>>

@@

Mainstream social media platforms can reduce the visibility of the content created or shared by specific users, whenever they violate the platforms' rules. The implementation of this policy varies across platforms and is not easy to verify ex-post. In what follows we provide means to verify this policy on Twitter and Facebook.

@@

% Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist\footnote{See \href{https://meta.wikimedia.org/wiki/Spam\_blacklist}{meta.wikimedia.org/wiki/Spam\_blacklist}.},
%CrowdTangle and Buzzsumo provide access to the interaction metrics on content (engagement), but not to the number of views (or reach), therefore we can only verify the {\it reduce} measure indirectly through a decrease in engagement. If Facebook was reducing the reach of a content via a lower ranking, we would expect the content to be less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless, they are explicitly condemning the material.\footnote{See the article in Wired by Martineau (2020)~\cite{wiredalexjones} and see the section {\it So what happened with InfoWars? They were up on Friday and now they are down?} \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{about.fb.com/news/2018/08/enforcing-our-community-standards/}.} To verify the measure, we used the ``/posts/search'' endpoint\footnote{see the documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Search}{github.com/CrowdTangle/API/wiki/Search}.} of the CrowdTangle API, to collect $37$ $242$ Facebook public posts that had shared a URL link containing ``infowars.com'', published between January $1$, $2019$ and December $31$, $2020$.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the $27$ $721$ posts directly sharing an Infowars link.}

@@

The number of public posts sharing an $Infowars$ link remained globally stable throughout $2019$ (see figure \ref{infowars1} top panel). Thus the measure announced by Facebook doesn't seem to have prevented users from sharing an $Infowars$ link. Nevertheless, a clear drop in engagement was observed on May 2, 2019 (see figure \ref{infowars1} bottom panel). The number of reactions, shares and comments per post have decreased respectively  by $-94\%$,  $-96\%$ and $-93\%$  when comparing the two months before and after May 2, 2019. This suggests the measure taken by Facebook in May $2019$, is not a ban per se, but rather a $reduce$ measure. This is because users could still post $Infowars$ links, but these posts generated less engagement. It should be noted that the engagement metrics increased again by the end of $2019$ / beginning of $2020$, suggesting that the $reduce$ measure may have been lifted a few months after its implementation.

%As CrowdTangle is tracking posts only from certain public groups and pages, we also used the ``/search/articles'' endpoint of the Buzzsumo API, to gather a richer Facebook dataset. We collected the engagement data for the $14$ $232$ articles crawled by Buzzsumo from the $Infowars$ website between January $1$, $2019$ and December $31$, $2020$.  %\footnote{The command can be found: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_infowars\_data.py}{here}.} 
%We observe that the articles published after May $2$, $2019$ received less Facebook engagement than the ones published before (see figure \ref{infowars2}), with a percentage change of -$97\%$ for the reactions, -$59\%$ for the shares and -$97\%$ for the comments. An increase in engagement was also observed in $2020$. It reinforces the hypothesis that Facebook reduced the reach of posts sharing Infowars links only during a few months in $2019$.

%\begin{figure}[h]
%	\centering
%	\includegraphics[scale=0.35]{./img/infowars/fb_infowars_3.png}
%	\caption{Facebook engagement metrics (average number of reactions, shares and comments per article) %for the Infowars articles published in $2019-2020$ and gathered from the Buzzsumo API. The red line marks %the date of May $2$, $2019$, when Facebook has announced the ban regarding Infowars.}
%	\label{infowars2}
%\end{figure}

%We found irregularities in the number of Infowars articles collected from Buzzsumo. While Infowars usually publishes around 800 articles per month, only 53 articles were collected between June 11 to July 11, 2020\footnote{In Figure \ref{infowars2} the engagement metrics were only shown for days in which at least 5 Infowars articles were collected from Buzzsumo, to ensure reliable results.}. A temporary crawling problem coming from Buzzsumo may have caused this lack of data. Because no database is perfect, we would like to highlight the importance of cross-checking information between different sources when possible.

Another measure that platforms can apply is to prevent users from sharing specific types of content, in this example URLs coming from a specific domain name. The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} Facebook has announced on December $20$, $2019$ that {\it The BL is now banned from Facebook} for coordinated inauthentic behavior (see Facebook Newsroom (2019)~\cite{newsroom1}), which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. Coordinated inauthentic behavior is a distinct phenomenon from disinformation according to Facebook, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false”.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.} Nevertheless, for the case of the Beauty of Life, both misinformation and coordinated inauthentic behavior were attested, according to the fact-checking organization Snopes which had reported about The BL’s activity to Facebook\footnote{\href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}}.

To verify Facebook’s ban of The BL domain name, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from the Buzzsumo API\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.}. We used the ``/search/articles'' endpoint to collect the engagement metrics of the $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.

@@

The number of Facebook reactions, shares and comments dropped to zero for TheBL’s articles published after December $1$, $2019$ (see figure \ref{fb_bl} top panel), indicating the start of the ban. We can note that although the ban was communicated in an article~\cite{newsroom1} published on December $20$, $2019$, it seems to have actually started on December $1$, $2019$. The communication around the ban appeared to have discouraged The Beauty of Life to proceed with their activity. Indeed the number of articles they published daily was around $50$ until December $20$, $2019$, when it decreased drastically to reach around $5$ to $10$ articles published per day (see figure \ref{fb_bl} bottom panel).

Using Buzzsumo data, we ascertained that links from thebl.com were not shared on Facebook anymore. The ban started on December $1$, $2019$, and appeared to be still enforced in June 2021.
